---
title: "LichenDu_HW2b"
author: "LichenDu"
date: "2/17/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

###2b Secant method

####Secant function
```{r}

#StartingValue needs to be two numbers
Secant=function(func,StartingValue,Tolerance,MaxNumberOfIterations){ 
  i=0 #number of iterations
  X1=StartingValue[1]
  X2=StartingValue[2]
  Y1=func(X1)[1]
  Y2=func(X2)[1]
  Deviation=1000
  allx=c()
  ally=c()
  allslope=c()
  allx[1]=X1
  allx[2]=X2
  ally[1]=Y1
  ally[2]=Y2
  allslope[1]=func(X1)[2]
  allslope[2]=func(X2)[2]

    while ((i<MaxNumberOfIterations)&&(Deviation>Tolerance))
  {
      
      if ((Y2=="NaN")||(Y1=="NaN")){
      cat("Function not defined error")
      cat("\n",Y2,Y1)
      break
          }
      
    # Find the next (X1,X2)-value using Newton-Raphson's formula
    if(Y1==Y2){
    NewX=X2-Y2*(X2-X1)/(Y2-Y1+Tolerance)
    NewY=func(NewX)[1]
    Deviation=abs(NewY)}
      else{
        NewX=X2-Y2*(X2-X1)/(Y2-Y1)
      NewY=func(NewX)[1]
      Deviation=abs(NewY)
      }

    #update value of x1 and x2
    X1=X2
    Y1=func(X1)[1]
    X2=NewX
    Y2=func(X2)[1]
    
    # increase the value of your iteration counter
    i=i+1
    cat(paste("\nIteration ",i,":   X=",NewX,"  Y=",NewY))
    allx[i+2]=NewX
    ally[i+2]=NewY
    allslope[i+2]=func(NewX)[2]
  
    }
    
  # output the result
  if (Deviation<Tolerance){
    cat(paste("\nFound the root point: ",NewX, " after ", i, "iterations"))
  }else{
    cat(paste("\nConvergence failure. Deviation: ",Deviation, "after ", i, 	"iterations"))}    
  
  # have the function return the answer
  df=cbind(allx,ally,allslope)
  return(df)
}

```


####Secant Plot function
```{r}
Secant_plot=function(func){
curve(func,lwd=5,xlim=c(min(df[,1]),max(df[,1])),ylim=c(min(df[,2]),max(df[,2]))) 
abline(h=0)
n=length(df[,1])
for (i in 1:(n-2)){
segments(df[i,1],0,df[i,1],df[i,2],lty=2,col="orange",lwd=2) #(x1,0) to (x1,y1)
segments(df[i,1],df[i,2],df[i+1,1],df[i+1,2],lty=2,col="red",lwd=2) #(x1,y1) to (x2,y2)
segments(df[i+1,1],df[i+1,2],df[i+2,1],0,lty=2,col="red",lwd=2) #(x2,y2) to (x3,0)
}
}
```


####Newton-Raphson function
```{r}
# Define your Newton-Raphson function  
NewtonRaphson<-function(func,StartingValue,Tolerance,MaxNumberOfIterations){
  i=0
  X=StartingValue
  Y=func(X)[1]
  Deviation=abs(Y)
  allx=c()
  ally=c()
  allslope=c()
  allx[1]=X
  ally[1]=Y
  allslope[1]=func(X)[2]
  while ((i<MaxNumberOfIterations)&&(Deviation>Tolerance))
  { Z=c()
    Z[1]=func(X)[1]
    Z[2]=func(X)[2]
    if ((Z[1]=="NaN")||(Z[2]=="NaN")){
      cat("\nFunction or derivative not defined error.\n")
      break
    }
   #update X and Y
   X=X-Z[1]/Z[2]
   Y=func(X)[1]
   Deviation<-abs(Y)
   i<-i+1
   allx[i+1]=X
   ally[i+1]=Y
   allslope[i+1]=func(X)[2]
   cat(paste("\nIteration ",i,":   X=",X,"  Y=",Y))
  }
  if (Deviation<Tolerance){
    cat(paste("\nFound the root point: ",X, " after ", i, "iterations"))
  }else{
    cat(paste("\nConvergence failure. Deviation: ",Deviation, "after ", i, 	"iterations"))}   
  df=cbind(allx,ally,allslope)
  return(df)
}

```

####Newton plot function
```{r}
Newton_plot=function(func){
curve(func,xlim=c(min(df[,1]),max(df[,1])),ylim=c(min(df[,2]),max(df[,2])),lwd=5) 
abline(h=0)
n=length(df[,1])
for (i in 1:(n-1)){
segments(df[i,1],0,df[i,1],df[i,2],lty=2,col="orange",lwd=2) #(x1,0) to (x1,y1)
segments(df[i,1],df[i,2],df[i+1,1],0,lty=2,col="red",lwd=2) #(x1,y1) to (x2,0)
}
}
```


#### 2 Functions
```{r}
#Functions
F1=function(x){
  return(c(cos(x)-x,-sin(x)-1))}
F11=function(x){
  return(cos(x)-x)}

F2=function(x){  
  return(c(log(x)-exp(-x),1/x+exp(-x)))}
F22=function(x){  
  return(log(x)-exp(-x))}

```


####Function1
```{r}
df=Secant(F1,c(1,2),0.0005,200)
Secant_plot(F11)

df=NewtonRaphson(F1,1,0.0005,200)
Newton_plot(F11)

df=NewtonRaphson(F1,2,0.0005,200)
Newton_plot(F11)
```



###Function2
```{r}
df=Secant(F2,c(1,2),0.0005,200)
Secant_plot(F22)

df=NewtonRaphson(F2,1,0.0005,200)
Newton_plot(F22)

df=NewtonRaphson(F2,2,0.0005,200)
Newton_plot(F22)
```


####Comparison: On those two functions, NewtonRaphson method has less iterations and performs better than Secant method. But when I try function1 in another startvalue, the situation changes.


####Function1 start from (1,10)
```{r}
df=Secant(F1,c(1,10),0.0005,200)
Secant_plot(F11)

df=NewtonRaphson(F1,1,0.0005,200)
Newton_plot(F11)

df=NewtonRaphson(F1,10,0.0005,200)
Newton_plot(F11)
```



####Conclusion: I cannot draw a conclustion regarding on the performance of these two root-finding methods cause their performance depends. Both starting values and function curve will affect the number of iterations.